\documentclass[../main.tex]{subfiles}


\begin{document}
\section{Motivation and Background}
\label{sec:motivation-and-background}

\subsection{Scaling operating system installation to large number of hosts}
The simplest way of installing an operating system on a computer is to use a bootable disk.
Bootable disk are removable media, such as USB flash drives or CD-ROMs, that contains software
which loads the operating system installer or standalone operating systems which does not require installation.

This method is simple and reliable, but it has some drawbacks. Primary, it requires physical access to the computer.
This is not a problem when installing an operating system on a personal computer, but it is a problem when installing
an operating system on remote computers especially when there are many of them. Another problem is that it is not
scalable, because it requires a person to manually install the operating system on each computer individually.

A solution is needed for environments which have to manage large number of computers such a data centers, businesses
or universities. The minimal requirements for such a solution are:
\begin{itemize}
  \item It should allow remote operating system installation. Without physical access to the computer
  \item It should be scalable. Adding new computers to the system should not require manual intervention and be manageable even for non-technical users
  \item The cost of maintaining the system should be low and should not require specialized hardware
  \item Virtualized system should be supported in the same manner as physical systems
\end{itemize}

\subsection{System maintenance and recovery}
Remote operating system installation is not the only problem that system administrators face. They are also
tased with maintaining the system and recovering it in a case of a failure. While most of the failures can be performed
remotely using Secure Shell (SSH) ore Remote Desktop Protocol (RDP), there are some failures that require physical access.
This type of failures are often caused by a failure in the booting process which will result in the operating system
not booting entirely or booting into a state which lacks the necessary tools to recover the system remotely.
These failures called for an on site recovery by a technician, which is both costly and time-consuming.

As long as the computer is able to properly communicate with the network there is another solution for this category
of failures. The solution involves booting computer into a specialized minimal operating system which provides
the necessary tools to recover the system remotely. This type of operating systems are called Live CD \cite{livecdlist} or Live USB.
They differ from the standard operating systems in that they do not require installation and can operate without
a persistent storage such as a hard drive or a solid state drive. They are loaded into the memory of the computer.

An exemplar of such an operating system is \texttt{SysRescueCD} \cite{sysrescuecd}.
Yet, according to the information available to the author \texttt{SysRescueCD} does not provide a way to remotely access the computer,
with its default configuration. This however can be achieved by using a tool called \texttt{OpenSSH} \cite{openssh} which is a free implementation of the Secure Shell protocol.
This configuration can be saved on Live CD's ram disk image which will be loaded after booting. In the case of
\texttt{SysRescueCD} the modifications should be made to \texttt{airootfs.sfs} file in the sysrescue ISO image.

To fulfil the promise of fully remote solution for system maintenance and recovery, the solution should be able to
boot a target computer into recovery system without physical access to the computer. Afterwards the recovery process
would be performed remotely using Secure Shell Protocol by an administrator.

\subsection{System maintenance messages}
Another valuable addition would be a mechanism for displaying messages of pending system maintenance the users of the system.
This feature is mostly targeted at universities and businesses where the user would be sitting in front of the target computer.
The message would inform the user that the system is going to be maintained due to a failure or a scheduled maintenance.
There already exists solutions for displaying such messages inside a running operating system, but according
to the author knowledge there is no solution for displaying scheduled maintenance messages on a computer which is in \texttt{Preboot Execution Environment}\cite{pxespec}.

By providing such a mechanism the system administrator would be able to inform the user of the target computer that
the system is going to be maintained even in a case of a failure which prevents the operating system from booting.
The only requirement for this mechanism to work is that the computer is able to boot into \texttt{Preboot Execution Environment}.
Which is a reasonable assumption, because the \texttt{PXE} is a part of the computer's network card firmware

\subsection{Dynamic operating system selection}


\end{document}